---
title: "Statistical disclosure control on EU-SILC microdata"
subtitle: "Risk, protection and information loss for a person-level file and a magnitude table"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    self_contained: true
params:
  output_dir: "../output"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE,
                      fig.width = 8, fig.height = 4.2, dpi = 144)
library(data.table)
library(ggplot2)
out <- function(f) fread(file.path(params$output_dir, f))
info <- jsonlite::read_json(file.path(params$output_dir, "run_info.json"))
risk <- out("risk.csv")
supp <- out("suppressions.csv")
ind <- out("indicators.csv")
ovl <- out("ci_overlap.csv")
tab <- out("table_protected.csv")
tab_sum <- out("table_summary.csv")
audit <- out("table_audit.csv")
audit0 <- out("table_audit_without_secondary.csv")

fmt <- function(x, d = 0) formatC(x, format = "f", digits = d, big.mark = ",")
pct <- function(x, d = 1) paste0(formatC(x, format = "f", digits = d), "%")

# Reference palette (validated categorical slots 1-3, neutral gray).
pal <- c(blue = "#2a78d6", orange = "#eb6834", aqua = "#1baf7a", gray = "#d9d8d4")
theme_report <- theme_minimal(base_size = 11) +
  theme(panel.grid.minor = element_blank(),
        panel.grid.major = element_line(colour = "#ecebe8", linewidth = 0.3),
        axis.title = element_text(colour = "#52514e"),
        axis.text = element_text(colour = "#52514e"),
        plot.title.position = "plot",
        legend.position = "top", legend.justification = "left")

orig <- risk[step == "no protection"]
final <- risk[step == "after local suppression"]
n_keys <- 8
scen_short <- function(x) sub(":.*", "", x)
```

<div class="alert alert-warning" role="note">
<strong>The data are synthetic.</strong> The EU-SILC public microdata are
simulated by Eurostat from models of the real survey: they contain no real
person, and Eurostat states that they cannot be used to draw conclusions about
the population. This project treats the file <em>as if</em> it were
confidential, to show the full disclosure-control workflow. All risk figures
and indicator values below illustrate the method; they say nothing about real
people or regions.
</div>

## Summary

This project plays the role of a statistical office that must release a
**public-use microdata file** and a **published table** from a household
survey without disclosing information about any respondent. It uses the
Italian file of the EU-SILC public microdata (`r info$year`,
`r fmt(info$persons)` persons aged 16 and over in `r fmt(info$households)`
households) as if it were the confidential source.

- In the original file, `r fmt(orig$sample_uniques)` records
  (`r pct(100 * orig$sample_uniques / orig$records)`) have a combination of
  eight publicly knowable characteristics that nobody else in the sample shares.
  The model-based estimate is `r fmt(orig$expected_reidentifications, 1)`
  expected re-identifications.
- Three protection scenarios were compared. All of them remove every sample
  unique. They differ in how much detail they sacrifice to get there.
- The recommended scenario, **`r final[scenario %like% "^S3", scenario]`**,
  leaves no combination shared by fewer than 5 records, reduces expected
  re-identifications to `r fmt(final[scenario %like% "^S3", expected_reidentifications], 2)`,
  and changes the headline income and poverty indicators by less than
  `r pct(ind[scenario %like% "^S3", max(abs(rel_diff_pct))], 2)`.
- In the table of employee income by region, occupation and sex,
  `r tab_sum$primary` cells are sensitive and `r tab_sum$secondary` more
  are hidden to stop them from being recomputed from the margins. Without
  the secondary suppressions, `r sum(!audit0$protected)` of the
  `r nrow(audit0)` sensitive cells could be recomputed exactly; with them, none can.

## Why disclosure control

European statistics are produced under a legal duty of statistical
confidentiality (Regulation (EC) No 223/2009, Article 20, and Principle 5 of
the European Statistics Code of Practice). Removing names and identifiers is
not enough: a person can be recognised from a combination of ordinary
characteristics, such as region, age, household size and occupation.
Statistical disclosure control (SDC) measures that risk and reduces it while
keeping the data useful.

## Data and scenario

**Source.** Eurostat, EU-SILC public microdata for Italy, file
`IT_PUF_EUSILC.zip`, SHA-256 `r substr(info$sha256, 1, 16)`…, downloaded from
the Eurostat website. The public microdata are **fully synthetic**: they are
simulated from models of the real survey and contain no real respondent.
Eurostat states that they cannot be used for inference about the population.
Here they are used for what they are meant for: training and method
development. The risk figures below describe this synthetic file, not real
people.

**Intruder scenario.** An intruder knows some facts about a person and looks
for records that match them. The facts considered are the *key variables*
(quasi-identifiers):

| Key variable | EU-SILC variable | Detail in the source |
|---|---|---|
| Region | DB040 | NUTS 2 (21 regions) |
| Sex | PB150 | 2 categories |
| Age | PX020 | single years, 16 to 80+ |
| Marital status | PB190 | 5 categories |
| Citizenship | PB220A | Italy / other EU / non-EU |
| Education | PE040 | ISCED 1997, 7 levels |
| Occupation | PL051 | ISCO-08 major group, 9 groups |
| Household size | HX040 | 1 to 10 |

Survey identifiers (PB030, PX030) are direct identifiers and are replaced by
random numbers. Employee income (PY010G) and household disposable income
(HY020) are confidential numerical variables. Weights (PB040) are kept.

## Risk in the original file

Two families of measures are used.

- **k-anonymity** counts records whose combination of keys appears fewer than
  *k* times in the sample. It needs no model and is easy to explain.
- **Individual risk** (Benedetti and Franconi, the method used by Istat and
  implemented in sdcMicro) estimates the probability that a match is correct,
  using the sampling weights to estimate how many people in the population
  share the same combination. The sum over records is the expected number of
  re-identifications. *Household risk* is the probability that at least one
  member of a household is re-identified, since identifying one member
  exposes the others.

```{r risk-original}
knitr::kable(
  data.table(
    Measure = c("Records", "Sample uniques", "Records violating 3-anonymity",
                "Records violating 5-anonymity", "Expected re-identifications",
                "Global risk", "Household risk"),
    Value = c(fmt(orig$records), fmt(orig$sample_uniques), fmt(orig$violating_3_anonymity),
              fmt(orig$violating_5_anonymity), fmt(orig$expected_reidentifications, 1),
              pct(orig$global_risk_pct, 2), pct(orig$household_risk_pct, 2))
  ), align = "lr")
```

The two families disagree, and the reason matters. About two thirds of the
records are unique in the sample, but each sampled person represents about 1,400
people (the mean weight), so the model expects few of those sample uniques to be
unique in the population as well. Agencies usually require both: a k-anonymity
threshold for the release file, and a check that the model-based risk is low.

## Protection methods

Each scenario applies three standard methods in this order (Hundepool et al.,
*Handbook on Statistical Disclosure Control*, 2012):

1. **Global recoding** of the key variables into broader categories.
2. **Local suppression**: individual key values are blanked, choosing the
   values that break the most risky combinations, until every combination is
   shared by at least *k* records. Region, sex and age are suppressed last
   because they matter most to users.
3. **Microaggregation** (MDAV, groups of 3) of the two income variables: each
   value is replaced by the mean of a group of three similar values, so no
   exact income is released. Household income is treated once per household,
   so all members of a household keep the same value.

```{r scenarios}
knitr::kable(data.table(
  Scenario = c("S1", "S2", "S3"),
  Region = c("NUTS 2", "NUTS 1", "NUTS 1"),
  Age = c("5-year bands", "10-year bands", "10-year bands"),
  `Other keys` = c("source detail", "grouped", "grouped"),
  k = c(3, 3, 5)
), align = "lllll")
```

"Grouped" means: occupation in four skill groups, education as low / medium /
high, marital status as never married / married / previously married,
household size top-coded at 5 or more.

## Risk after protection

```{r risk-steps}
r <- risk[step != "no protection",
          .(Scenario = scen_short(scenario), Step = step,
            `Sample uniques` = fmt(sample_uniques),
            `< 3-anonymity` = fmt(violating_3_anonymity),
            `< 5-anonymity` = fmt(violating_5_anonymity),
            `Expected re-identifications` = fmt(expected_reidentifications, 2),
            `Household risk` = pct(household_risk_pct, 3))]
knitr::kable(r, align = "llrrrrr")
```

Recoding alone does most of the work for S2 and S3. In S1, which keeps
detailed codes, recoding leaves many combinations rare, so local suppression
has to blank many more values to reach the same guarantee.

## Information loss

### Suppressed values

```{r supp-chart, fig.height = 3.8}
s <- copy(supp)[, scenario := scen_short(scenario)]
labels <- c(region = "Region", sex = "Sex", age = "Age", marital = "Marital status",
            citizenship = "Citizenship", isced = "Education", isco1 = "Occupation",
            hh_size = "Household size")
s[, variable := factor(labels[variable], levels = rev(labels))]
ggplot(s, aes(x = suppressed, y = variable, fill = scenario)) +
  geom_col(position = position_dodge(width = 0.8), width = 0.72) +
  scale_fill_manual(values = unname(pal[c("blue", "orange", "aqua")]), name = NULL) +
  scale_x_continuous(labels = scales::comma, expand = expansion(mult = c(0, 0.05))) +
  labs(x = "Values blanked by local suppression", y = NULL) +
  theme_report
```

```{r supp-table}
st <- supp[, .(`Values suppressed` = fmt(sum(suppressed)),
               `Share of key values` = pct(100 * sum(suppressed) / (orig$records * n_keys), 2)),
           by = .(Scenario = scen_short(scenario))]
knitr::kable(st, align = "lrr")
```

### Indicators

Each indicator is computed on the original and on the protected file with
the same code. The poverty threshold is 60% of the national median
equivalised disposable income, as in the Eurostat indicator `ilc_li02`.
Because the file is synthetic, the levels are not real Italian statistics;
only the differences matter. The differences come almost entirely from the
microaggregation of incomes, which is the same in all scenarios: suppressing
key values hardly moves these indicators. The scenarios differ instead in
the detail they publish and in the number of blanked values.

```{r indicators}
iw <- dcast(ind[, .(indicator, domain, scenario = scen_short(scenario), rel_diff_pct)],
            indicator + domain ~ scenario, value.var = "rel_diff_pct")
base <- unique(ind[, .(indicator, domain, value_original)])
iw <- merge(base, iw, by = c("indicator", "domain"))
iw[, value_original := ifelse(indicator %like% "AROP|Gini", fmt(value_original, 3), fmt(value_original))]
for (s in intersect(c("S1", "S2", "S3"), names(iw))) set(iw, j = s, value = pct(iw[[s]], 2))
setnames(iw, c("indicator", "domain", "value_original"), c("Indicator", "Domain", "Original value"))
knitr::kable(iw, align = "llrrrr",
             caption = "Relative difference, protected vs original")
```

### Regression

A user fitting a wage equation (log employee income on sex, age band and
NUTS 1 area, weighted) should reach the same conclusions on the protected
file. The confidence-interval overlap (Karr et al., 2006) is 1 when the 95%
intervals coincide and 0 when they do not overlap at all.

```{r overlap}
ov <- ovl[term != "(Intercept)", .(`Mean overlap` = round(mean(overlap), 2),
                                   `Lowest overlap` = round(min(overlap), 2),
                                   `Coefficients` = .N),
          by = .(Scenario = scen_short(scenario))]
knitr::kable(ov, align = "lrrr")
```

## Risk-utility trade-off

```{r ru-map, fig.height = 4}
ru <- rbind(
  data.table(label = "Original", loss = 0, er = orig$expected_reidentifications),
  merge(final[, .(scenario, er = expected_reidentifications)],
        supp[, .(loss = 100 * sum(suppressed) / (orig$records * n_keys)), by = scenario],
        by = "scenario")[, .(label = scen_short(scenario), loss, er)]
)
ggplot(ru, aes(x = loss, y = er)) +
  geom_point(size = 3.2, colour = pal[["blue"]]) +
  geom_text(aes(label = label), nudge_x = 0.25, hjust = 0, size = 3.6, colour = "#0b0b0b") +
  scale_y_log10() +
  scale_x_continuous(expand = expansion(mult = c(0.05, 0.25))) +
  labs(x = "Key values suppressed (% of all key values)",
       y = "Expected re-identifications (log scale)") +
  theme_report
```

Points lower down are safer; points further left lose less detail. Global
recoding is not captured on the horizontal axis: S2 and S3 publish coarser
categories than S1 for every record.

**Recommendation.** S3 is the release candidate. It reaches 5-anonymity with
far fewer suppressions than S1 because recoding removed the rare categories
first, and its indicators are practically unchanged. Moving from S2 to S3
(k = 3 to k = 5) roughly doubles the suppressions but removes the remaining
records that share their combination with only 3 or 4 others. S1 keeps regional
detail but must blank many more values, mostly occupation, and offers only
3-anonymity. Users who need NUTS 2 detail should use the scientific-use file
under a research contract rather than a more detailed public file.

## Table protection

Published tables leak information too. The table is the estimated total
employee income by region (Italy, NUTS 1, NUTS 2), occupation (ISCO major
group) and sex: `r fmt(tab_sum$cells)` cells including all margins.

- **Primary suppression.** A cell is sensitive if it has fewer than 3
  contributors (threshold rule) or if the second-largest contributor could
  estimate the largest within 10% by subtracting its own value from the total
  (p% rule, p = 10).
- **Secondary suppression.** A suppressed cell could be recomputed from
  its row or column totals, so additional cells are hidden. sdcTable's
  `SIMPLEHEURISTIC` chooses them across the region hierarchy.
- **Audit.** For every sensitive cell, a linear program computes the tightest
  interval an intruder can derive from the published cells. The cell is
  protected only if that interval is not a single value.

```{r table-summary}
knitr::kable(data.table(
  Item = c("Cells", "Published", "Primary suppressions", "Secondary suppressions",
           "Income in suppressed detailed cells",
           "Sensitive cells recomputable without secondary suppression",
           "Sensitive cells recomputable with secondary suppression"),
  Value = c(fmt(tab_sum$cells), fmt(tab_sum$published), fmt(tab_sum$primary),
            fmt(tab_sum$secondary), pct(tab_sum$income_hidden_pct, 1),
            paste(sum(!audit0$protected), "of", nrow(audit0)),
            paste(sum(!audit$protected), "of", nrow(audit)))
), align = "lr")
```

```{r table-pattern, fig.height = 6.2}
p <- tab[nchar(region) == 4 & occupation != "Total" & sex != "Total"]
p[, status := fcase(sdcStatus == "u", "Primary suppression",
                    sdcStatus == "x", "Secondary suppression",
                    default = "Published")]
p[, status := factor(status, levels = c("Published", "Primary suppression", "Secondary suppression"))]
p[, sex := fifelse(sex == "F", "Women", "Men")]
p[, occupation := sub("^OC0", "", occupation)]
p[occupation == "1", occupation := "0-1"]
p[, region := factor(region, levels = rev(sort(unique(region))))]
ggplot(p, aes(x = occupation, y = region, fill = status)) +
  geom_tile(colour = "white", linewidth = 1) +
  facet_wrap(~sex) +
  scale_fill_manual(values = c(`Published` = pal[["gray"]],
                               `Primary suppression` = pal[["orange"]],
                               `Secondary suppression` = pal[["blue"]]),
                    name = NULL, drop = FALSE) +
  labs(x = "Occupation (ISCO-08 major group)", y = "NUTS 2 region") +
  theme_report + theme(panel.grid.major = element_blank())
```

`r sum(p$status == "Primary suppression" & p$sex == "Women")` of the
`r sum(p$status == "Primary suppression")` sensitive detailed cells are for
women: occupations with few female employees (skilled agricultural workers,
group 6; plant and machine operators, group 8; managers, groups 0-1) in small
or southern regions. Secondary suppressions fall on the matching cells for
men and on neighbouring regions, so that no row or column total can be used
to recover them.

## Limitations

- The source is synthetic, so the risk levels are illustrative. On a real
  confidential file the same code would give different numbers.
- The intruder scenario fixes eight key variables. A different scenario (for
  example an intruder who knows income) needs its own key variables.
- Microaggregation by itself does not guarantee that a very high income
  cannot be linked to an outlier; a production release would add top-coding
  of income checked against the actual distribution.
- `SIMPLEHEURISTIC` does not minimise the information hidden. An
  optimisation method (`OPT`, or τ-ARGUS) would usually hide less.

## Reproducibility

Source file pinned by SHA-256 in `data/manifest.json`; sdcMicro
`r info$sdcMicro`, sdcTable `r info$sdcTable`; run at `r info$run_at`.
Code: <https://github.com/simones99/eu-silc-sdc>.
